### 2009-07-02 paper.tex
fixed some typos

A singlet transducer (which emits, but does not absorb, symbols) lies at the root of the phylogeny and 
serves as a generative model of the ancestral sequence.
To represent the evolution of an ancestral sequence into many descendant sequences,
we place a branch transducer on each branch of the phylogeny.
Figures \ref{fig:parent} and \ref{fig:threeway} show simple examples
of building a model to represent such an evolution.

Throughout this paper we frequently refer to two and
three-taxon (star) phylogenies.  In all cases, the sequence $W$ is
assumed to be the (unobserved) ancestral sequence and the sequences
$X$, $Y$, and $Z$ the (observed) extant sequences (as in Figures
\ref{fig:parent} and \ref{fig:threeway}).

@@

We first consider the {\em perfect alignment rate}; that is, the number of times each method gets the alignment exactly correct.
Theory predicts that Maximum Likelihood inference, using the correct model and parameters, should be the algorithm that scores the most perfect estimates.
Inspecting \figref{perfect}, we find this to be the case, except where the outgroup is very distant and the bins are probably undersampled
(i.e. the departure from prediction that we observe for low-identity alignments is not statistically significant:
when the perfect alignment rate drops below $\sim 5/125$, then 125 trials are probably too few).
We also note that the ML version of \stemloc\ is near-optimal, despite the \stemloc\ pair-SCFG being
slightly different from the TKFST pair-SCFG in parameterization and structure
 (e.g. \stemloc's grammar does not allow insertion/deletion of entire substructures).
The ML version of \stemloc\ is also observed to have a slightly higher perfect alignment rate than the posterior-decoding version (\stemloc-AMA).
Finally, we note that the structure-blind models (TKF91 and Long Indel) perform consistently worse than the structure-aware methods;
furthermore, both linear (TKF91) and affine (Long Indel) gap-penalties perform equally bad in this test
(note that the TKFST model, from which the true alignments were simulated, does not allow long-indel events,
which may partly explain why affine gap-penalties do not help in this benchmark).

>>>>

A singlet transducer (which emits, but does not absorb, symbols) lies at the root of the phylogeny and 
serves as a generative model of the ancestral sequence.
To represent the evolution of an ancestral sequence into many descendant sequences,
we place a branch transducer on each branch of the phylogeny.
% Figures \ref{fig:parent} and \ref{fig:threeway} show simple examples
% of building a model to represent such an evolution.

Throughout this paper we frequently refer to two and
three-taxon (star) phylogenies.  In all cases, the sequence $W$ is
assumed to be the (unobserved) ancestral sequence and the sequences
$X$, $Y$, and $Z$ the (observed) extant sequences.
% (as in Figures \ref{fig:parent} and \ref{fig:threeway}).

@@

We first consider the {\em perfect alignment rate}; that is, the number of times each method gets the alignment exactly correct.
Theory predicts that Maximum Likelihood inference, using the correct model and parameters, should be asymptotically optimal (if one only counts perfect guesses).
Inspecting \figref{perfect}, we find this to be almost the case; the exception is when the outgroup is very distant and the bins may be undersampled
(the departure from prediction that we observe for low-identity alignments is not statistically significant:
when the optimal success rate drops below $\sim 5/125$, then 125 trials are probably insufficient to compare two near-optimal methods).
We also note that the ML version of \stemloc\ is near-optimal, despite the \stemloc\ pair-SCFG being
slightly different from the TKFST pair-SCFG in parameterization and structure
 (e.g. \stemloc's grammar does not allow insertion/deletion of entire substructures).
The ML version of \stemloc\ is also observed to have a slightly higher perfect alignment rate than the posterior-decoding version (\stemloc-AMA).
Finally, we note that the structure-blind models (TKF91 and Long Indel) perform consistently worse than the structure-aware methods;
furthermore, both linear (TKF91) and affine (Long Indel) gap-penalties perform equally bad in this test
(note that the TKFST model, from which the true alignments were simulated, does not allow long-indel events,
which may partly explain why affine gap-penalties do not help in this benchmark).
